peter_models

### 2024-12-14 RQ1_Plot.qmd
add OR plot code

```{r}
load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
```

@@

rename_map <- c(
    "Q271_reordered(2) Very important" = "Racial Belonging - Very Important",
    "Q271_reordered(1) Extremely important" = "Racial Belonging - Extremely Important",
    "S3B(2) Woman" = "Gender - Woman",
    "S3B(3) Non-binary" = "Gender - Non-binary",
    "race_recodedBlack" = "Race - Black",
    "race_recodedAsian American" = "Race - Asian",
    "race_recodedLatino" = "Race - Latino",
    "race_recodedAIAN" = "Race - Native",
    "race_recodedNHPI" = "Race - NHPI",
    "age_recoded50-59" = "Age - 50-59",
    "age_recoded60-69" = "Age - 60-69",
    "age_recoded70+" = "Age - 70+",
    "educSome college experience" = "Education - Some College",
    "educCollege graduate" = "Education - College Graduate",
    "Q21(2) Democrat" = "Party ID - Democrat",
    "Q21(3) Independent" = "Party ID - Independent",
    "Q21(4) Other party" = "Party ID - Other",
    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "Interaction: Racial Belonging & Gender: Woman"
  )

@@

>>>>

```{r}
load("/Users/jenniferli/Desktop/2020cmps_clean2.RData")
```

```{r}
```

```{r}
summary(dataset_clean)
```

@@

rename_map <- c(
    "Q271_reordered(2) Very important" = "Racial Belonging - Very Important",
    "Q271_reordered(1) Extremely important" = "Racial Belonging - Extremely Important",
    "S3B(2) Woman" = "Gender - Woman",
    "S3B(3) Non-binary" = "Gender - Non-binary",
    "race_recodedBlack" = "Race - Black",
    "race_recodedAsian American" = "Race - Asian",
    "race_recodedLatino" = "Race - Latino",
    "race_recodedAIAN" = "Race - Native", # why use Native instead of AIAN?
    "race_recodedNHPI" = "Race - NHPI",
    "age_recoded50-59" = "Age - 50-59",
    "age_recoded60-69" = "Age - 60-69",
    "age_recoded70+" = "Age - 70+",
    "educSome college experience" = "Education - Some College",
    "educCollege graduate" = "Education - College Graduate",
    "Q21(2) Democrat" = "Party ID - Democrat",
    "Q21(3) Independent" = "Party ID - Independent",
    "Q21(4) Other party" = "Party ID - Other",
    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "Interaction: Racial Belonging & Gender: Woman"
  )

@@

# Plot with Andrea's following comments all considered:

1.  make sure to use appropriate variable labels.
2.  I see that you ordered them based on the OR values, but I think it’s confusing to see the responses to the questions in different places. I would still group the questions together, and maybe put them in descending order within each question?
3.  Also make sure the reference levels are clear so that the reader can appropriately interpret the plot

```{r}
odds_ratio_plot_labelled <- function(model) {
  # Get odds ratios and CIs
  or_data <- exp(cbind(OR=coef(model), confint(model)))

  # Create a mapping for variable names with grouping prefixes
  rename_map <- c(
    # Education Group
    "educCollege graduate" = "1. Education: College Graduate",
    "educSome college experience" = "1. Education: Some College",

    # Gender Group
    "S3B(2) Woman" = "2. Gender: Woman",
    "S3B(3) Non-binary" = "2. Gender: Non-binary",

    # Age Group
    "age_recoded70+" = "3. Age: 70+",
    "age_recoded60-69" = "3. Age: 60-69",
    "age_recoded50-59" = "3. Age: 50-59",

    # Race Group
    "race_recodedBlack" = "4. Race: Black",
    "race_recodedAsian American" = "4. Race: Asian",
    "race_recodedLatino" = "4. Race: Latino",
    "race_recodedAIAN" = "4. Race: Native",
    "race_recodedNHPI" = "4. Race: NHPI",

    # Racial Belonging Group
    "Q271_reordered(2) Very important" = "5. Racial Belonging: Very Important",
    "Q271_reordered(1) Extremely important" = "5. Racial Belonging: Extremely Important",

    # Party ID Group
    "Q21(2) Democrat" = "6. Party ID: Democrat",
    "Q21(3) Independent" = "6. Party ID: Independent",
    "Q21(4) Other party" = "6. Party ID: Other",

    # Interactions
    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "7. Interaction: Racial Belonging & Gender: Woman"
  )

  or_data_sig_renamed <- or_data_sig
  rownames(or_data_sig_renamed) <- rename_map[rownames(or_data_sig_renamed)]

  # Custom sorting function
  sort_by_group_and_name <- function(x) {
    # Split the strings to get group numbers and names
    groups <- as.numeric(sub("^(\\d+)\\..+$", "\\1", x))
    # Get the part after the colon for sorting within groups
    names <- sub("^\\d+\\.[^:]+: (.+) \\(ref.+$", "\\1", x)
    # Create a data frame for sorting
    sort_df <- data.frame(
      original = x,
      group = groups,
      name = names,
      stringsAsFactors = FALSE
    )
    # Sort by group (ascending) and then by name (descending)
    sort_df <- sort_df[order(sort_df$group, -xtfrm(sort_df$name)), ]
    return(sort_df$original)
  }

  # Apply the custom sorting
  sorted_names <- sort_by_group_and_name(rownames(or_data_sig_renamed))
  or_data_sig_renamed <- or_data_sig_renamed[sorted_names, ]

  ggplot() +
    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
    geom_point(aes(x = or_data_sig_renamed[,1], 
                   y = factor(rownames(or_data_sig_renamed), 
                          levels = rev(rownames(or_data_sig_renamed)))), 
              size = 4, color = "blue") +
    geom_errorbarh(aes(xmin = or_data_sig[,2], 
                       xmax = or_data_sig[,3],
                       y = factor(rownames(or_data_sig_renamed), 
                              levels = rev(rownames(or_data_sig_renamed)))),
                   height = 0.3, color = "black") +
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 6, color = "black"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      axis.title.x = element_text(size = 12, face = "bold"),
      axis.title.y = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.margin = margin(1, 1, 1, 1, "cm"),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.minor.y = element_blank()
    ) +
    labs(
      title = "Significant Variables with Confidence Intervals",
      x = "Odds Ratio",
      y = "Variables"
    )
}

Note: Some variable levels are excluded due to filtering for significant variables. To view the reference level for each variable group, please see the below table.

## Reference Levels Table

| Variable Group | Reference Level |
|----|----|
| 1\. Education | No college experience |
| 2\. Gender | Man |
| 3\. Age | 18-29 years old |
| 4\. Race | White |
| 5\. Racial Belonging | Not At All Important |
| 6\. Party ID | Republican |
| 7\. Interaction: Racial Belonging & Gender: Woman | Man, Not At All Important |
